\section{Conclusion}
\label{sec:conclusion}

We introduced \emph{semantic influence}---a novel extension to taint analysis that tracks how user-controlled data propagates through LLM API calls to reach dangerous sinks. Unlike traditional taint analysis, which loses track at API boundaries, our model captures the empirical reality that LLM outputs remain influenced by their inputs even though they are syntactically independent.

Our formalization introduces controllability (separated from impact severity) with position-dependent decay (0.80 for prompts, 0.70 for context, 0.50 for system messages) and multi-hop attenuation. We implemented this model in a practical static analyzer and evaluated it on real-world LLM applications.

\paragraph{Key Results.}
\begin{itemize}
    \item \textbf{76\% precision, 60\% recall} on the OWASP LLM vulnerability testbed
    \item Zero false positives on security commit benchmark (positive-only corpus; see Section~\ref{sec:evaluation} for caveats)
    \item \textbf{1,816 taint-based findings} across 5 popular LLM applications, 94\% classified as requiring semantic influence (based on tool output; see Section~\ref{sec:case-studies} for validation limitations)
    \item \textbf{2 vulnerabilities} reported to maintainers via pull requests
\end{itemize}

\paragraph{Limitations and Future Work.}
Our current implementation performs single-file analysis only; extending to cross-file interprocedural analysis would improve applicability to production LLM services where data flows span modules. The influence decay parameters are empirically chosen based on our testbed; a more principled approach might learn these from larger labeled datasets. Our evaluation benchmarks are relatively small (73 + 75 samples) and heavily positive-only; validation on larger mixed corpora would strengthen claims. Finally, while we focus on Python, the semantic influence model applies to any language with LLM integrations.

\paragraph{Broader Impact.}
As LLM-integrated applications become ubiquitous, the attack surface for prompt injection expands. Semantic influence analysis provides a foundation for securing this new class of applications by extending proven static analysis techniques to handle the unique challenge of LLM transformations.

Our tool is available at \url{https://github.com/deosha/aisentry} and evaluation artifacts at \url{https://github.com/deosha/llm-sec-eval}.
